The not so widening achievement gap: an interntational comparison of the evolution of the SES achievement gap from 2000-2015

### 2017-08-30 paper/paper.Rnw
Created graph that shows all gaps and the modeling - two models left

difference_data <-
  complete_gaps %>%
  filter(type_test == "math", type == "90th/10th SES gap") %>%
  select(wave, country, difference) %>%
  rename(year = wave) %>%
  mutate(year = as.character(year))

@@

ready_data <-
  difference_data %>%
  left_join(tracking_data, by = c("country" = "cntry_name")) %>%
  left_join(central_examination) %>%
  mutate(num_tracks = ifelse(tracks15y == 1, 1, 0) %>% as.factor,
         age_selection = selage) %>%
  filter(!is.na(num_tracks), !is.na(age_selection), !is.na(difference))

ready_data_age <- mutate(ready_data, age_selection = ifelse(selage >= 15, 1, 0) %>% as.factor)

@@

mod_tracking

@@

bayes_R2(mod_tracking)
marginal_effects(mod_tracking,
                 effects = "ztrack:zvoc",
                 int_conditions = list(zvoc = quantile))

@@

pp_check(mod_tracking, type = "dens_overlay", nsamples = 100)
pp_check(mod_tracking, type = "ecdf_overlay", nsamples = 10)
pp_check(mod_tracking, type = "boxplot", nsamples = 30)
pp_check(mod_tracking, type = "stat", nsamples = 200) # very nice to compare central tendency

pp_check(mod_tracking, type = "error_scatter_avg", nsamples = 1000)

@@

# update all of this for the previous model!
y_rep <- posterior_predict(mod_tracking)

replication <-
  map(1:2000, ~ {
  ready_data_two %>%
  mutate(pred_sample = y_rep[sample(1:4000, 1), , drop = TRUE]) %>%
  summarize(perc_sample = mean(pred_sample > cum_diff),
            minus_sample = mean(pred_sample - cum_diff))
})

y_pred_data <-
  replication %>%
  enframe() %>%
  unnest()

y_pred_data %>% qplot(perc_sample, data = ., geom = "histogram", bins = 50)

y_pred_data %>% {quantile(.$perc_sample)}

y_pred_data %>%
  summarize(mean_perc = mean(perc_sample < 0.5))

@@

m_data %>%
  filter(ratio_pred < 1) %>%
  ggplot(aes(reorder(country, -ratio_pred), 1 - ratio_pred)) +
  geom_col() +
  scale_y_continuous(breaks = seq(0, 0.30, 0.05), limits = c(0, 0.3)) +
  coord_flip()

>>>>

@@

dif_data <- function(my_data, tracking, central, which_gap = "90th/10th SES gap") {
  my_data %>%
    filter(type_test == "math", type == which_gap) %>%
    select(wave, country, difference) %>%
    rename(year = wave) %>%
    mutate(year = as.character(year)) %>%
    left_join(tracking_data, by = c("country" = "cntry_name")) %>%
    left_join(central_examination) %>%
    mutate(num_tracks = ifelse(tracks15y == 1, 1, 0) %>% as.factor,
           age_selection = selage) %>%
    filter(!is.na(num_tracks), !is.na(age_selection), !is.na(difference))

}

ready_data <- dif_data(complete_gaps, tracking_data, central_examination)

ready_data_age <-
  dif_data(complete_gaps, tracking_data, central_examination) %>%
  mutate(age_selection = ifelse(selage >= 15, 1, 0) %>% as.factor)

gaps <- c("90th/10th SES gap", "80th/20th SES gap", "70th/30th SES gap")

all_gaps_models <-
  map(gaps, function(gap) {
    ready_data_age <-
      dif_data(complete_gaps, tracking_data, central_examination, which_gap = gap) %>%
      mutate(age_selection = ifelse(selage >= 15, 1, 0) %>% as.factor)

    mod_tracking
})

model_data_merged <-
  map(all_gaps_models, ~ fixef(.x)[, c(1, 3, 4)] %>% as_tibble) %>%
  setNames(gaps) %>%
  enframe() %>%
  unnest(value) %>%
  mutate_if(is_double, round, 2) %>%
  transmute(model = name,
           terms = all_gaps_models[[1]] %>% fixef %>% row.names %>% rep(3),
           estimate = Estimate,
           lower = `2.5%ile`,
           upper = `97.5%ile`)

model_data_merged %>%
  ggplot(aes(terms, estimate, colour = model)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1) +
  geom_point() +
  geom_hline(yintercept = 0) +
  scale_colour_discrete(name = NULL) +
  scale_y_continuous(limits = c(-0.5, 2)) +
  ggtitle("Coef from linear model of inequality on degree of tracking and vocational tracking") +
  coord_flip()

@@

plot_interaction <-
  marginal_effects(mod_tracking,
                 effects = "ztrack:zvoc",
                 int_conditions = list(zvoc = quantile))

@@

@@

pp_check(mod_tracking, type = "dens_overlay", nsamples = 100)
pp_check(mod_tracking, type = "ecdf_overlay", nsamples = 10)
pp_check(mod_tracking, type = "boxplot", nsamples = 30)
pp_check(mod_tracking, type = "stat", nsamples = 200) # very nice to compare central tendency
pp_check(mod_tracking, type = "error_scatter_avg", nsamples = 1000)

@@

# update all of this for the previous model!
# y_rep <- posterior_predict(mod_tracking)
# 
# replication <-
#   map(1:2000, ~ {
#   ready_data_two %>%
#   mutate(pred_sample = y_rep[sample(1:4000, 1), , drop = TRUE]) %>%
#   summarize(perc_sample = mean(pred_sample > cum_diff),
#             minus_sample = mean(pred_sample - cum_diff))
# })
# 
# y_pred_data <-
#   replication %>%
#   enframe() %>%
#   unnest()
# 
# y_pred_data %>% qplot(perc_sample, data = ., geom = "histogram", bins = 50)
# 
# y_pred_data %>% {quantile(.$perc_sample)}
# 
# y_pred_data %>%
#   summarize(mean_perc = mean(perc_sample < 0.5))

@@

m_data %>%
  ggplot(aes(reorder(country, -ratio_pred), 1 - ratio_pred)) +
  geom_col() +
  scale_y_continuous(breaks = seq(-0.3, 0.3, 0.5), limits = c(-0.3, 0.3)) +
  coord_flip()
